\documentclass{article}

\usepackage{amsmath}
\usepackage{enumitem}
\usepackage{amsfonts}
\newcommand{\Mod}[1]{\ (\mathrm{mod}\ #1)}

\title{CS 2051 Practice Exam 2}
\author {TAs: Alan Tao, Nithya Jayakumar, Rishi Raman}
\date{Spring 2022}

\begin{document}
    \maketitle
\section{Modular Arithmetic}

\begin{enumerate}[label=\alph*.]
\item Find the last digit of $7^{100}$
\item $5x \equiv6$ (mod 8)
\item Solve for x. $\sum_{n = 1} ^ {k} n! \equiv x$ (mod 9)
\item Prove that $2^{n} + 6 \cdot 9^{n} \equiv 0 $ (mod 7) for any natural number $n$. Hint: Try induction ;) 
\item Find the last 3 digits of $2027 ^ {2027}$
\end{enumerate}

\section{Euclidean Algorithm}
Work backwards to find Bezout coefficients for the $g.c.d.$.
\begin{enumerate}
    \item Find $g.c.d.(12, 18)$.
    \item Find $g.c.d.(111, 201)$.
    \item Find $g.c.d.(123, 277)$.
\end{enumerate}
\section{Binomial Theorem}
    \begin{enumerate}
        \item What is the coefficient of $x^{12}y^{13}$ found in $(x + y)^{25}$? What about $(2x - 3y)^{25}$?
        \item Give a formula for for the coefficient of $x^k$ in the expansion of $(x^2 - \frac{1}{x})^{100}$, where k
            is an integer.
        \item Prove $\sum _{k = 0}^n \binom{n}{k} = 2^n.$
    \end{enumerate}

\section{Chinese Remainder Theorem}
\begin{enumerate}
    \item Find x such that $2x \equiv 5 \Mod{7}$ and $3x \equiv 4 \Mod{8}.$
    \item Find x such that $x \equiv 1 \Mod{3},$ $x \equiv 4 \Mod{5},$ and $x \equiv 6 \Mod{7}.$
    \item Find all x such that $x \equiv 2 \Mod{3},$ $x \equiv 1 \Mod{4},$ $x \equiv 3 \Mod{5}.$
\end{enumerate}

\section{Time Complexity}

\begin{enumerate}[label=\alph*.]

\item Show that $f(x) = x^{2} + 3x - 2$ is $O(x^{3})$
\item Show that $h(x) = (x+1)^{2}log(x^4 - 3) + 2x^{3}$ is $O(x^{3})$

\end{enumerate}

\section{aaaand lots of Combo}
\begin{enumerate}
    \item A coin is flipped 10 times where each flip comes up either heads or tails. How many possible outcomes 1) are there in total? 2) contain exactly two heads? 3) contain at most three tails? 4) contain the same number of heads and tails?
    \item A club has 25 members. 1) How many ways are there to choose four members of the club to serve on an executive committee? 2) How many ways are there to choose a president, vice president, secretary, and treasurer of the club, where no person can hold more than one office?
    \item Suppose that a department contains 10 men and 15 women. How many ways are there to form a committee with six members if it must have the same number of men and women?
    \item How many non-negative solutions are there to the equation $x_1 + x_2 + x_3 + x_4 + x_5 = 21$ if 1) $x_1 \geq 1$? 2) all $x_i \geq 2$? 3) $0 \geq x_1 \geq 10$? 4) $0 \geq x_1 \geq 3$, $1 \geq x_2 \geq 4$, and $x_3 \geq 15$?
    \item How many ways are there to distribute six indistinguishable balls into nine distinguishable bins?
    \item How many positive integers less than 1,000,000 have the sum of their digits equal to 19?
    \item How many different strings can be made from the letters in ABRACADABRA, using all the letters?
    \item How many ways are there to travel in xyz space from the origin (0, 0, 0) to the point (4, 3, 5) by taking steps one unit in the positive x direction, one unit in the positive y direction, or one unit in the positive z direction? (Moving in the negative x, y, or z direction is prohibited, so that no backtracking is allowed.)
\end{enumerate}

\end{document}
