\documentclass{article}
\usepackage{url}
\usepackage{hyperref}
\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    filecolor=magenta,      
    urlcolor=blue,
    pdftitle={Overleaf Example},
    pdfpagemode=FullScreen,
    }

\usepackage{indentfirst}

\urlstyle{same}
\title{CS 2051 Syllabus}
\author{Spring '23}
\date{\today}

\begin{document}
\maketitle

\section{Instructor Information}


\noindent Gerandy Brito: \url{gbrito3@gatech.edu}

\noindent TAs

\noindent Sarthak Mohanty: \url{smohanty@gatech.edu}

\noindent Nithya Jayakumar: \url{n.jayakumar@gatech.edu}

\section{Office Hours}
Will be posted in a Canvas announcement soon.


\section{Welcome and Format}

Lecture Schedule: Tuesday, Thursday 8 - 9:15 AM

Location: CCB 016

Lectures are synchronous and on campus. We will implement quizzes on ZyBooks to account for participation and you are responsible for those.
\begin{itemize}
    \item We do not record our lectures.
    \item All information for the class will be shared on Canvas and during lecture. 
    \item Canvas modules will be used to host lecture content and homework assignments. These modules will be updated twice weekly (within 24 hours after each class).  The most recent modules will be at the top of the list.
    Exams will be held during class and cannot be taken asynchronously without an official ODS accommodation.
    \item Homework will be assigned on Canvas and submitted on Gradescope. You may type or hand-write your homework. Make sure it is legible and easy to read. You may be penalized otherwise.
    \item Ed Discussion is a class forum that you can use to chat about any questions you're having on the lectures.  This should not be used to ask for direct solutions to homework problems. 
        \begin{itemize}
        \item Please keep all general posts open to the public as other students may benefit and may help you faster than we can. 
        \item Private questions that directly apply to your solution of a homework problem should be marked private.
        \end{itemize}
\end{itemize}

\section{Textbooks}
Zybooks Links to an external site: We will use Zybooks for graded participation quizzes. To access it you must:
\begin{itemize}
    \item Click any zyBooks assignment link in your learning management system: this ensures your canvas account is directly linked to zybooks.  Do not make a separate Zybooks account as you will have syncing issues.
    
    \item Subscribe: the book costs \$58.
\end{itemize}
 

Discrete Mathematics and Its Applications by Kenneth Rosen, 8th edition. Lectures will follow along with the 8th edition of this book and it is expected that you will read the corresponding sections.  The format of the book is irrelevant (i.e.: hardcopy, kindle edition, etc) but if you use an older edition that won’t be an excuse to miss homework due to a different enumeration of the problems and layout of the sections. There is a solution guide for the odd numbered problems that is not required but could be helpful.  While this book isn't required for any homework assignments, I highly suggest utilizing it.

ISBN10: 125967651X
ISBN13: 9781259676512

\section{Schedule}
See living schedule on Canvas.

\section{Course Objectives}

Help students gain an understanding of basic primitives and paradigms of the mathematical theory of computation.
Students will learn to write formal proofs and understand logic. Students will take an in-depth look at various forms
of induction and recursion. Various areas of mathematics are explored to help provide a foundation for computation.
The course provides a basis for understanding and developing clear logic, understanding simple encryption
techniques, writing regular expressions, designing computational models and more.


\section{Topic List}
Propositions, Propositional Logic

Propositional Equivalences

Quantified Expressions, Nested Quantifiers, Rules of Inference

Introduction to Proofs including Direct Proof, Proof by Contradiction, Proof by Contraposition, Proof by Counterexample

Induction & Recursion, including Proofs by Mathematical Induction, Strong Induction

Recurrence Equations, Recursively Defined Sets, Recursively Defined Functions, Recursively Defined Strings,
Recursively Defined Sequences, Recursively Defined Algorithms

Sets, Functions, Sequences, Summations, Strings as Sequences

Introduction to Algorithms including Searching, Sorting

Growth Rates of Functions - Big O, Time Complexity, Space Complexity

Algorithm Pseudocode, Algorithmic Paradigms including Divide & Conquer Algorithms, Greedy Algorithms,
Brute-Force Algorithms, Recursive Algorithms

Number Theory: Divisibility, Modular Arithmetic, Division Algorithm, Euclid’s Algorithm, Primality

Cryptography: RSA and Secure Computing 

Basic Counting (Sum Rule, Product Rule, Inclusion/Exclusion), Pigeonhole Principle, Permutations, Combinations,
& Finite Probability

Models of Computation: Deterministic Finite Automata, Regular Expressions

Application of techniques to domains of relevance to computer science


\section{Grading Scheme}
After all your grades are entered letter grades are computed according to the following brackets:

A 90-100

B 80-89.99

C 70-79.99

D 60-69.99

F 0-59.99

These brackets may change, but only in your benefit. Rounding and curving is unlikely.

Grade Breakdown:

\begin{itemize}
    \item 10\% Zybooks
    \item 60\% Exams (4)
    \item 10\% Project
    \item 20\% HWs
\end{itemize}

We will soon be providing an Excel Sheet you can input your grades into for each assignment into to generate your current final grade.

\section{Zybooks}
Every Tuesday we will release a set of quizzes on ZyBooks related to the class content that you must answer for participation credit. These quizzes will be cumulative for the entire week's material.  You will have exactly 7 days to complete that week's quizzes.
\begin{itemize}
    \item Attending and watching every lecture should be more than enough to prepare you for these quizzes.
    \item No extensions of any kind will be offered.
    \item Please note that no sections will be dropped. Zybooks ensures that you're doing the required readings.
\end{itemize}


\section{Homeworks}

Homework will be posted on Canvas, and submitted on Gradescope. You are encouraged to type your homework. If you choose not to, please be considerate of your TAs and write clear and legible. We will penalize your score if your homework is not legible.  The lowest homework will be dropped.

Regrade Requests: Once grades are released you will have one week to dispute your grade. Regrade requests must be based on solid arguments and frivolous requests won’t be addressed. Examples of frivolous requests are “I feel I deserve more points”; “I think the grading was too harsh”; “Please check my solution again”; “I do not know where I made a mistake. Can you read my solution, and tell me what I did wrong, and if I deserve more points?”  Regrade requests will be handled through GradeScope.

Late submissions will be allowed up to 2 days late.  Your first late submission will incur no late penalty.  Remember that you must turn in your assignments on time.  If you are even a second late, then the assignment is considered late.  Please don't wait until the last minute to submit as you may have technical difficulties.  We cannot grant exemptions for this.  All other late submissions will result in the following deductions:

\begin{itemize}

    \item Early (greater than 24 hours before due date): +2.5 points to that assignment
    \item On Time: No penalty
    \item 1 day late: 10 point automatic penalty
    \item 2 days late: 25 point automatic penalty
\end{itemize}

Collaboration policy: Collaboration is allowed. You must write your own solutions though. On every homework, you should list the names of those students you collaborated with. Remember that copying a solution from your peer or from the internet is plagiarism and is penalized by the GT Code of Conduct. If cheating is detected, a report will be filed with the Office of Student Integrity which can lead to a 0 on the assignment or potentially an F in the course overall.

\section{Exams}
Exams will be conducted in class. By registering to take this class, you are responsible for taking the exam during this time. There will be no make up exam unless you have an official excuse from Student Services. If that is the case, make sure to contact the staff at least 48 hours in advance. There are obvious exceptions to this rule (like an accident the day of the exam...which hopefully won’t happen). Final decision is at the sole discretion of the instructors.

The regrade policy for exam is one week and must be done by making a regrade request on GradeScope.

Personal calculators are not allowed on exams. No notes allowed.

No collaboration allowed on Exams. Remember that copying a solution from your peer or from the internet is plagiarism and is penalized by the GT Code of Conduct. If cheating is detected, a report will be filed with the Office of Student Integrity which can lead to a 0 on the assignment or potentially an F in the course overall.

\section{Project}
2051 students will be working on a final group project throughout the semester. This project will include several intermediate milestones as well as a final product and presentation at the end of the semester. There will be four graded parts to the project: A thesis/outline, a first draft, a final product, and a presentation. Some information:
\begin{itemize}
    \item Groups will be three students each. Work on finding people to form a team with as groups will be finalized in week 2. We will be releasing a group sign up document soon.
    \item We will release some topic idea options for the project which students can use (first come, first serve for signup) however coming up with your own topic ideas is encouraged. Groups that come up with their own ideas will have to get the topic approved.
    \item The final product (and subsequently first draft) will be in the form of a paper. Details will come soon.
    \item The final presentation will be a science fair-style affair, with groups setting up posters or presentations in a room with the intention of presenting their findings and ideas to their peers and instructors.
    \item Grading breakdown for project:
        \begin{itemize}
            \item 1\%: Thesis/outline
            \item 1\%: First draft
            \item 4\%: Final paper
            \item 4\%: Presentation
        \end{itemize}
\end{itemize}

\section{Students with Special Accommodations}
Classroom accommodations for students with documented disabilities will be made, if necessary. These accommodations must be arranged in advance and in accordance with the Office of Disability Services: http://disabilityservices.gatech.edu

\section{Dean of Students Office, CARE Center, Counseling Center, Stamps Health Services, and the Student Center}
The \href{https://mentalhealth.gatech.edu/}{CARE Center} and the \href{https://mentalhealth.gatech.edu/}{Counseling Center}, Stamps Health Services, and the Dean of Students Office will offer both in-person and virtual appointments. Student Center services and operations are available on the \href{https://studentcenter.gatech.edu/}{Student Center} website. For more information on these and other student services, contact the Dean of Students or the \href{https://studentlife.gatech.edu/}{Division of Student Life}.

The following guidance addresses practices that were common during AY2020-2021 and may still be relevant to some courses for Fall 2021, such as online courses. If applicable to a given course, this information may be inserted into the syllabus.  

\section{COVID-19}
Students are expected to be familiar with and abide by the Institute guidelines, information, and updates related to Covid-19. Find campus operational updates, Frequently Asked Questions, and details on campus surveillance testing and vaccine appointments on the \href{https://health.gatech.edu/coronavirus}{Tech Moving Forward site}.    

\section{Recordings of Class Sessions and Required Permissions}
Classes may not be recorded by students without the express consent of the instructor unless it is pursuant to an accommodation granted by the Office of Disability services. Class recordings, lectures, presentations, and other materials posted on Canvas are for the sole purpose of educating the students currently enrolled in the course. 

Students may not record or share the materials or recordings, including screen capturing or automated bots, unless the instructor gives permission. Digitally proctored exams may require students to engage the video camera, but those recordings will not be shared with or disclosed to others without consent unless legally permitted.    
\begin{itemize}
    \item For classes where participation is voluntary, students who participate with their camera engaged or utilize a profile image are agreeing to have their video or image recorded. 

    \item  For classes requiring class participation, if students are identifiable by their names, facial images, voices, and/ or comments, written consent must be obtained before sharing the recording with persons outside of currently enrolled students in the class. 
\end{itemize}


\end{document}
